---
title: "Week 1 Quiz — Getting and Cleanning data"
output: html_document
---

Question 1
----------

The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv 

and load the data into R. The code book, describing the variable names is here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf 

How many properties are worth $1,000,000 or more?

``` {r} 
data <- read.csv("getdata-data-ss06hid.csv",header=TRUE,sep=",")
newdata <- subset(data, VAL >= 24)
total <- length(newdata[,newdata$VAL])
print(total) 
```


Question 2
----------

Use the data you loaded from Question 1. Consider the variable FES in the code book. Which of the "tidy data" principles does this variable violate?

Question 3
----------

Download the Excel spreadsheet on Natural Gas Aquisition Program here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx

Read rows 18-23 and columns 7-15 into R and assign the result to a variable called:
 
 ```dat```
 
What is the value of:

``` sum(dat$Zip*dat$Ext,na.rm=T) ```

(original data source: http://catalog.data.gov/dataset/natural-gas-acquisition-program)

``` {r}
library(xlsx)
colIndex <- 7:15
rowIndex <- 18:23
dat <- read.xlsx("getdata-data-DATA.gov_NGAP.xlsx",sheetIndex=1,colIndex=colIndex,rowIndex=rowIndex)
sum(dat$Zip*dat$Ext,na.rm=T)
```


Question 4
----------

Read the XML data on Baltimore restaurants from here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml 

How many restaurants have zipcode 21231?

``` {r}
library(XML)
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml"
download.file(fileUrl,destfile = "restaurants.xml",method="curl")
doc <- xmlTreeParse("restaurants.xml",useInternal=TRUE)
rootNode <- xmlRoot(doc)
ziplist <- as.numeric(xpathSApply(rootNode,"//zipcode",xmlValue)) #find all nodes called 'zipcode'
length(ziplist[ziplist==21231])
```


Question 5
----------

The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here: 

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv 

using the fread() command load the data into an R object

```DT ```

Which of the following is the fastest way to calculate the average value of the variable

``` pwgtp15 ```

broken down by sex using the data.table package?

```{r}
library(data.table)
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
download.file(fileUrl, destfile = "idaho.csv", method="curl")
dateDownloaded <- date()
print(dateDownloaded)
DT <- fread("idaho.csv")
file.info("idaho.csv")$size
system.time(mean(DT$pwgtp15,by=DT$SEX))
system.time(sapply(split(DT$pwgtp15,DT$SEX),mean))
system.time(tapply(DT$pwgtp15,DT$SEX,mean))
system.time(DT[,mean(pwgtp15),by=SEX])
system.time(mean(DT[DT$SEX==1,]$pwgtp15)) + system.time(mean(DT[DT$SEX==2,]$pwgtp15))
#system.time(rowMeans(DT)[DT$SEX==1]) + system.time(rowMeans(DT)[DT$SEX==2])
```

